### 2008-09-02 doc/client/client_technical.tex
new structure

This document describes from a technical viewpoint a plugin based client library for the new 
Web Service (WS) based Advanced Resource Connector (ARC) middleware. The library is capable of 
handling resource discovery, job submission and job control and interoperates with several 
different grid flavours. The support for different grid flavours is enabled through the modular 
design using plugins specialized for each supported middleware, and presently supported are ARC0 
(ARC classic), ARC1 and the gLite middlewares. Future extensions to support additional middlewares 
involve plugin compilation only i.e. no recompilation of main libraries or client.

@@

Finally the library has been designed with care for later usage within a graphical user interface. This 
particularly applies to the job management part.

\chapter{Resource Discovery and Information Retrieval}
\label{sec:TargetDiscovery}
Almost every day a new grid cluster is purchased and made available around the world (either to the 
general public or restricted to authorized users). More available clusters create a need for better and 
faster resource discovery and information retrieval, making this task an utmost important component of the 
new client library.

The new client library's resource discovery and information retrieval component consists of three classes, one 
of which a base class for further grid middleware specific specialization (plugin):

\begin{itemize}
\item{{\bf TargetGenerator} is responsible for loading TargetRetriever plugins in accordance with the input URLs. 
e.g. if an URL pointing to an ARC1 resource is given the TargetRetrieverARC1 is loaded. When queried for resources 
(``targets'') the TargetGenerator passes the request to the loaded TargetRetrievers running them as individual threads 
for improved performance. The TargetGenerator keeps records of by TargetRetriever found servers and services 
(index or computing) in order to avoid multiple queries. Accepted targets are stored in the FoundTargets array kept by 
the TargetGenerator.}

\item{{\bf TargetRetriever} is base class for TargetRetriever grid middleware specializations. Contains the pure 
virtual method GetTargets which is to be implemented by the specialized class. While it is not mandatory it is recommended 
that the specialized class divides this function into two components: QueryIndex and InterrogateTarget. The former is 
responsible for querying index servers, and if the result yields an addresss to a different index server then the 
specialized TargetRetriever should call itself recursively with the found url. Discovered computing services are sent 
to the InterrogateTarget function (or if the TargetRetriever was initialized with a computing service url then it should 
jump directly to this point) which collects the detailed information about the service.}

\item{{\bf ExecutionTarget} is the class representation of a computing resource (queue) capable of running a grid job. It
serves as input to the broker which is foreseen to be able to select between different ExecutionTargets from different 
grid flavours without apriori knowing their difference. The ExecutionTarget class mimics the glue2 information model (with 
a flattened structure), and thus a mapping between attributes from other information systems into the glue2 format is needed. 
Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares. All information can be printed by 
invoking the ExecutionTarget's ``Stat'' command which prints all variables (if they have been filled during the information 
retrieval process).

Following a broker decision jobs are to be submitted. Since all information about the selected computation resides within the 
selected ExecutionTarget, the ExecutionTarget is capable of returning a submitter (see next section) capable of submitting a 
job to the resource which is represents.}
\end{itemize}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.75}{\includegraphics{TargetDiscovery.png}}}
\caption{\label{fig:ResDisc}Diagram depicting the process of resource discovery.} }
\end{figure}

@@

\chapter{Job Submission and Management}
\label{sec:JobManagement}
\section{Submission}
Job submission starts with the resource discovery and target preparation as outlined in the section above. Not before a list 
of possible targets (which allows the user) is available is the job description read in order to enable bulk job submission of 
widely different jobs without having to reperform the resource discovery. The job submission makes use of the following classes:

\begin{itemize}
\item{{\bf Broker} is not yet implemented and thus not described here.}

\item{{\bf JobDescription} is responsible for handling all job descriptions. Thus it is capable of handling the xRSL, JDL and 
JSDL job description languages and to some extent also translate between them. Validation (syntax and semantical) is performed 
automatically when the source is set. Features easy retrieval of list of local input files which should be uploaded to the 
computing service.}

\item{{\bf Submitter} is base class for grid specific specializations (plugin).  It submits job(s) to a service and 
uploads (by the job needed) local input files. Returns an XML node containing all needed information about the job 
for later job management. The Submitter is returned by the ExecutionTarget selected for job execution and thus the 
ExecutionTarget populates (through XML config element) the Submitter with information about submission endpoint (URL) 
and job description languages ``spoken'' by the target.}

\end{itemize}

@@

\section{Management}
Information about active jobs is stored in a local xml file which is used for later job management. As this file may contain 
information about jobs running on completely different grid flavours, job management should be handled using plugins similar to 
resource discovery and job submission. The following classes are used:

\begin{itemize}
\item{{\bf JobSupervisor} is responsible for loading the appropriate JobController for managing a job running on a specific 
grid flavour. In order to match a job identification string (jobid) with a grid flavour a lookup in the local xml file is 
performed. The JobSupervisor main functionality is to return a list of JobControllers which are capable of managing all jobs 
addressed by a certain command.}

\item{{\bf JobController} is both base class for grid specific specializations, but also the implementor of all public functionality 
offered by the JobControllers. In other words all virtual functions of the JobController are private. Upon initialization the 
JobController reads into and fills a local store of jobs called JobsStore.
The public functions of the 
JobController offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations for issueing the command. 
This ensures consistency in how the commands are handled }

\item{{\bf Job} is }
\end{itemize}

\chapter{Data Management}
\label{sec:DataManagement}

\chapter{Specialized Classes (Grid Flavour Plugins)}
\label{sec:plugins}
\section{ARC0 Plugins}
\section{ARC1 Plugins}
\section{gLite Plugins}

>>>>

This document describes from a technical viewpoint a plugin based client library for the new 
Web Service (WS) based Advanced Resource Connector (ARC) middleware. The library consists of 
a set of C++ classes for

\begin{itemize}
\item{handling proxy, user and host certificates,}
\item{performing resource discovery and information retrieval,}
\item{job submission and management and}
\item{data handling.}
\end{itemize}

All capabilities are enabled for three different grid flavours (Production ARC, ARC1 and gLite) 
through the modular design using plugins specialized for each supported middleware. Future 
extensions to support additional middlewares involve plugin compilation only i.e. no recompilation 
of main libraries or client.

@@

In the following we will give a functionality overview in Section~\ref{sec:FuncOver} while all technical 
details will be given in Section~\ref{sec:Implementation}. Section~\ref{sec:cli} will through examples 
show how command line interfaces can be built upon the library.

\chapter{Functionality Overview}
\label{sec:FuncOver}
The new ARClib makes extensive use of plugins for command handling. These plugins are handled by a set 
of higher level classes which thus are the ones to offer the plugin functionality to external calls. In 
this section an overview of the library's main functionality is given which also introduces the most 
important higher level classes.
\section{Resource Discovery and Information Retrieval}
\label{sec:TargetDiscovery}
With the increasing number of grid clusters around the world, a reliable and fast resource discovery and 
information retrieval capability is of crucial importance for a user interface. The new ARClib resource 
discovery and information retrieval component consists of three classes; The TargetGenerator, the 
TargetRetriever and the ExecutionTarget. Of these the TargetRetriever is a base class for further grid 
middleware specific specialization (plugin).

@@

\begin{figure}[ht]
\centering{{\scalebox{0.75}{\includegraphics{TargetDiscovery.png}}}
\caption{\label{fig:ResDisc}Diagram depicting the process of resource discovery and information retrieval.} }
\end{figure}

@@

\section{Job Submission}
\label{sec:JobSubmission}
Job submission starts with the resource discovery and target preparation as outlined in the section above. Not before a list 
of possible targets (which allows the user) is available is the job description read in order to enable bulk job submission of 
widely different jobs without having to reperform the resource discovery. The job submission makes use of the following classes:

@@

\section{Job Management}
Information about active jobs is stored in a local xml file which is used for later job management. As this file may contain 
information about jobs running on completely different grid flavours, job management should be handled using plugins similar to 
resource discovery and job submission. The following classes are used:

\section{Data Handling}
\label{sec:DataHandling}

\chapter{Implementation}
\label{sec:Implementation}
Clever text here.
\section{Generic Classes}
\subsection{TargetGenerator} The TargetGenerator is responsible for loading TargetRetriever plugins in accordance with the 
input URLs. e.g. if an URL pointing to an ARC1 resource is given the TargetRetrieverARC1 is loaded. When queried for resources 
(``targets'') the TargetGenerator passes the request to the loaded TargetRetrievers running them as individual threads 
for improved performance. The TargetGenerator keeps records of by TargetRetriever found servers and services 
(index or computing) in order to avoid multiple queries. Accepted targets are stored in the FoundTargets array kept by 
the TargetGenerator.

\subsection{TargetRetriever} The TargetRetriever is base class for TargetRetriever grid middleware specializations. Contains the pure 
virtual method GetTargets which is to be implemented by the specialized class. While it is not mandatory it is recommended 
that the specialized class divides this function into two components: QueryIndex and InterrogateTarget. The former is 
responsible for querying index servers, and if the result yields an addresss to a different index server then the 
specialized TargetRetriever should call itself recursively with the found url. Discovered computing services are sent 
to the InterrogateTarget function (or if the TargetRetriever was initialized with a computing service url then it should 
jump directly to this point) which collects the detailed information about the service.

\subsection{ExecutionTarget} The ExecutionTarget is the class representation of a computing resource (queue) capable of running a grid job. It
serves as input to the broker which is foreseen to be able to select between different ExecutionTargets from different 
grid flavours without apriori knowing their difference. The ExecutionTarget class mimics the glue2 information model (with 
a flattened structure), and thus a mapping between attributes from other information systems into the glue2 format is needed. 
Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares. All information can be printed by 
invoking the ExecutionTarget's ``Stat'' command which prints all variables (if they have been filled during the information 
retrieval process).

Following a broker decision jobs are to be submitted. Since all information about the selected computation resides within the 
selected ExecutionTarget, the ExecutionTarget is capable of returning a submitter (see next section) capable of submitting a 
job to the resource which is represents.

\subsection{Broker} is not yet implemented and thus not described here.

\subsection{JobDescription} The JobDescription is responsible for handling all job descriptions. Thus it is capable of handling the xRSL, JDL and 
JSDL job description languages and to some extent also translate between them. Validation (syntax and semantical) is performed 
automatically when the source is set. Features easy retrieval of list of local input files which should be uploaded to the 
computing service.

\subsection{Submitter} Submitter is base class for grid specific specializations (plugin).  It submits job(s) to a service and 
uploads (by the job needed) local input files. Returns an XML node containing all needed information about the job 
for later job management. The Submitter is returned by the ExecutionTarget selected for job execution and thus the 
ExecutionTarget populates (through XML config element) the Submitter with information about submission endpoint (URL) 
and job description languages ``spoken'' by the target.

\subsection{JobSupervisor} is responsible for loading the appropriate JobController for managing a job running on a specific 
grid flavour. In order to match a job identification string (jobid) with a grid flavour a lookup in the local xml file is 
performed. The JobSupervisor main functionality is to return a list of JobControllers which are capable of managing all jobs 
addressed by a certain command.

\subsection{JobController} is both base class for grid specific specializations, but also the implementor of all public functionality 
offered by the JobControllers. In other words all virtual functions of the JobController are private. Upon initialization the 
JobController converts the information stored in the local XML to Jobs which are stored in a JobControler internal store called JobsStore. 
The public functions of the JobController offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations 
for issueing the command. This ensures consistency in how the commands are handled between the different grid flavours. Jobs that are 
permanently removed following a command are removed from both the JobStore and local xml file. Thus it is foreseen that the JobControllers 
may be used in a future graphical user interface where it becomes a persistent object continously monitoring jobs.

\subsection{Job} is generic job class for storing all job related information. Attributes are derived from the glue2 information model 
and thus a mapping is needed for non glue2 compliant grid middlewares. Appendix XX shows the present mapping schema.

\subsection{UserConfig}

\section{Specialized Classes (Grid Flavour Plugins)}
\label{sec:plugins}
\subsection{ARC0 Plugins}
\subsection{ARC1 Plugins}
\subsection{gLite Plugins}

\section{Support Classes}
\subsection{URL}
\subsection{Logger}
